### 2024-02-07 README.Rmd
correct sourcing of image from vignette/articles folder, labelling of joined constituencies

---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  message = FALSE, 
  warning = FALSE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# sfislands

The goal of `sfislands` is to make it easier to deal with geographic datasets which contain islands. It does so using a tidy framework in the spirit of Josiah Parry's [sfdep](https://sfdep.josiahparry.com/) package.

-   These do not have to be "literal" islands but any situation where discontiguous geographical units are present.

Such a situation can lead to two issues.

-   Firstly, if unaddressed, the presence of such islands or exclaves can make certain types of contiguity-based modelling impossible.

-   Secondly, just because two areas are separated by, say, a body of water, this does not necessarily mean that they are to be considered independent of each other.

This package offers solutions to allow for the inclusion or exclusion of these units within an uncomplicated workflow.

## Installation

You can install the development version of `sfislands` from [GitHub](https://github.com/horankev/sfislands) with:

``` r
# install.packages("devtools")
devtools::install_github("horankev/sfislands")
```

## Summary of features

1.  The initial setting up neighbourhood structures can be frustrating for people who are eager to get started with fitting spatial models. This is especially so when the presence of discontiguities within a geographical dataset means that, even having set up a neighbours list, the model will still not run without further awkward data manipulations.

2.  As an aid to setting up neighbourhood structures, particularly when islands are involved, the package has a function to quickly map any neighbourhood structure for visual inspection. This can also be used to examine the output of `sfdep` neighbour functions. Such maps can be used to check if the structure makes sense, given the researcher's knowledge about the geography of the study area.

3.  If there are some neighbours assigned which are not appropriate, or if you wish to add additional ones, there are functions to allow this to be done in a straightforward and openly reportable way.

4.  Once an appropriate neighbourhood structure is in place, different types of statistical tests and models can be performed. `sfdep` contains functionality to perform such test, and the output from `sfislands` can be used in its functions.

5.  The contiguity outputs from `sfislands` can be directly used to fit different types of (multilevel) (I)CAR models using, for example, the `mgcv`, `brms`, `stan` or `INLA` packages.

6.  For `mgcv` in particular, the predictions of such models can be quite tedious to extract and visualise. `sfislands` can streamline this workflow from the human side. Furthermore, there is a function to draw maps of these predictions for quick inspection.

<img src="man/figures/logo.png" align="right" height="240"/>

## Functions overview

The following is a framework within which the `sfislands` functions could be used.

The first group can be seen as pre-functions, designed to create a neighbourhood structure suitable for certain types of models.

Once a model has been fit (using `mgcv` in this case), the post-functions can be used to extract the predictions.

### Step 1: Set up data ("*pre-functions*")

| function:               | purpose:                                                                                                                                   |
|-----------------|-------------------------------------------------------|
| **st_bridges()**        | *Create a neighbours list, matrix, or `sf` dataframe containing a neighbours list or matrix as column "nb", while accounting for islands.* |
| **st_quickmap_nb()**    | *Check contiguities visually on map.*                                                                                                      |
| **st_check_islands()**  | *Check assignment of island contiguities in a dataframe.*                                                                                  |
| **st_manual_join_nb()** | *Make manual changes to any connections.*                                                                                                  |
| **st_manual_cut_nb()**  | *Make manual changes to any connections.*                                                                                                  |

### Step 2: Create model

*Use the output of **st_bridges()** as both the data and neighbourhood inputs for a model using, for example, `mgcv`, `brms` or `inla`.*

### Step 3: Examine output ("*post functions*")

| function:               | purpose:                                                 |
|------------------|------------------------------------------------------|
| **st_augment()**        | *Augment the original dataframe with model predictions.* |
| **st_quickmap_preds()** | *Generate quick maps of these predictions.*              |

```{r, echo=FALSE}

## Data and packages

# Read in data and necessary packages.

# install.packages("devtools")
# devtools::install_github("horankev/sfislands")

# required packages
packages <- c(
  "tidyverse",
  "sf",
  "mgcv",
  "ggpubr",
  "spdep",
  "sfdep",
  "here"
)

# Install packages not yet installed
installed_packages <- packages %in% rownames(installed.packages())
if (any(installed_packages == FALSE)) {
  install.packages(packages[!installed_packages])
}

# Packages loading
invisible(lapply(packages, library, character.only = TRUE, quietly = TRUE))

library(sfislands)

```

## Examples

In the context of the constituencies of England, Scotland and Wales, `st_bridges` will define a neighbourhood structure which includes the islands. `st_quickmap_nb` will show a representation of this on a map:

```{r, echo=FALSE}
uk_election$country <- case_when(uk_election$region == "Wales" ~ "Wales",
                                  uk_election$region == "Scotland" ~ "Scotland",
                                  TRUE ~ "England")
```

```{r, fig.width=12, fig.height=10}
nbsf <- st_bridges(df = uk_election,
                   geom_col_name = "constituency_name",
                   link_islands_k = 2)
st_quickmap_nb(nbsf,
                    pointsize=0.05, 
                    title = "st_bridges() contiguities",
                    subtitle = "islands linked to k=2 nearest constituencies")

```

Neighbours can be manually added or removed using `st_manual_join_nb()` and `st_manual_cut_nb()`.

```{r}
st_bridges(df = uk_election|> filter(region %in% c("Wales","South West")),
           geom_col_name = "constituency_name",
           link_islands_k = 2)  |> 
  st_manual_join_nb("Gower","St Ives") |> 
  st_quickmap_nb() + 
  geom_sf_label(data=uk_election|> filter(constituency_name %in% c("Gower","St Ives")),
                aes(label=constituency_name),nudge_x = -30000)
```

## Modelling & post-functions

Having set up a neighbourhood structure and embedded it as a named list or matrix within the original `sf` dataset as column `nb`, there are some functions to make it easy to quickly perform ICAR smoothing, augment the original dataframe with these predictions, and visualise them.

For example, we can use the `mgcv` package to generate a Markov Random Field ICAR smooth of poor health across the study area. This is done very quickly by using `st_bridges()` to prepare the data, putting that inside the `mgcv` GAM formulation, and then piping into the `st_augment()` and `st_quickmap_preds()` functions.

```{r, fig.width=12, fig.height=6}

prep_data3 <- st_bridges(uk_election, "constituency_name") # decide upon the contiguities and add them to the df

model <- gam(con_swing ~ 
               s(region, bs="re") + # region level random intercept
               s(county, bs="re") + # county level random intercept
               s(county, degree_educated, bs="re") + # county level random coefficient
               s(constituency_name, bs='mrf', 
                 xt=list(nb=prep_data3$nb),k=10) + # ICAR constituency ICAR varying coefficients
               s(constituency_name, by=white, bs='mrf', 
                 xt=list(nb=prep_data3$nb),k=10), # ICAR constituency ICAR varying coefficients
             data=prep_data3, method="REML") |> 
  st_augment(prep_data3) |> # pipe into function to get estimates
  st_quickmap_preds() # pipe into this for visualisation

ggarrange(plotlist = model, legend = "none", nrow=1)
```

Further information and worked examples are available in the this [article](https://horankev.github.io/sfislands/).