### 2024-03-14 README.Rmd
improved readme

---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  message = FALSE, 
  warning = FALSE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

<img src="man/figures/logo.png" align="right" height="240"/>

# sfislands

The goal of `sfislands` is to make it easier to deal with geographic datasets which contain islands. It does so using a tidy framework in the spirit of Josiah Parry's [sfdep](https://sfdep.josiahparry.com/) package.

-   These do not have to be *islands* by a traditional *land and water* geographic definition, but any situation where discontiguous geographical units are present.

Such a situation can lead to two issues.

-   Firstly, if unaddressed, the presence of such islands or exclaves can make certain types of contiguity-based modelling impossible.

-   Secondly, just because two areas are separated by, say, a body of water, this does not necessarily mean that they are to be considered independent of each other.

This package offers solutions to allow for the inclusion or exclusion of these units within an uncomplicated workflow.

## Installation

Install the released version from CRAN

``` r
install.packages("sfislands")
```

You can install the development version of `sfislands` from [GitHub](https://github.com/horankev/sfislands) with:

``` r
# install.packages("devtools")
devtools::install_github("horankev/sfislands")
```

## Summary of features

1.  The initial setting-up of neighbourhood structures can be frustrating for people who are eager to get started with fitting spatial models. This is especially so when the presence of discontiguities within a geographical dataset means that, even having set up a neighbours list, the model will still not run without further awkward data manipulations.

2.  As an aid to setting up neighbourhood structures, particularly when islands are involved, the package has a function to quickly map any neighbourhood structure for visual inspection. This can also be used to examine the output of `sfdep` neighbour functions. Such maps can be used to check if the structure makes sense, given the researcher's knowledge about the geography of the study area.

3.  If some units have been assigned as neighbours inappropriately, or if we wish to add additional connections, there are functions to allow this to be done in a straightforward and openly reportable way.

4.  Once an appropriate neighbourhood structure is in place, different types of statistical tests and models can be performed. `sfdep` contains functionality to perform such tests, and the output from `sfislands` can be used in its functions.

5.  The contiguity outputs from `sfislands` can be directly used to fit different types of (multilevel) (I)CAR models using, for example, the `mgcv`, `brms`, `stan` or `INLA` packages.

6.  For `mgcv` in particular, the predictions of such models can be quite tedious to extract and visualise. `sfislands` can streamline this workflow from the human side. Furthermore, there is a function to draw maps of these predictions for quick inspection.

## Functions overview

The following is a brief explanations of the functions contained in the `sfislands` package and a workflow within which they could be used.

The first group can be seen as pre-functions, designed to create a neighbourhood structure suitable for certain types of models.

Once a model has been fit (using `mgcv` in this case), the post-functions can be used to extract the predictions.

### Step 1: Set up data ("*pre-functions*")

| function:               | purpose:                                                                                                                                   |
|-----------------|-------------------------------------------------------|
| **st_bridges()**        | *Create a neighbours list, matrix, or `sf` dataframe containing a neighbours list or matrix as column "nb", while accounting for islands.* |
| **st_quickmap_nb()**    | *Check contiguities visually on map.*                                                                                                      |
| **st_check_islands()**  | *Check assignment of island contiguities in a dataframe.*                                                                                  |
| **st_manual_join_nb()** | *Make manual changes to any connections.*                                                                                                  |
| **st_manual_cut_nb()**  | *Make manual changes to any connections.*                                                                                                  |

### Step 2: Create model

Use the output of `st_bridges()` as both the data and neighbourhood inputs for a model using, for example, `mgcv`, `brms` or `inla`.

### Step 3: Examine output ("*post functions*")

| function:               | purpose:                                                 |
|------------------|------------------------------------------------------|
| **st_augment()**        | *Augment the original dataframe with model predictions.* |
| **st_quickmap_preds()** | *Generate quick maps of these predictions.*              |

```{r, echo=FALSE}

## Data and packages

# Read in data and necessary packages.

# install.packages("devtools")
# devtools::install_github("horankev/sfislands")

# required packages
packages <- c(
  "tidyverse",
  "sf",
  "mgcv",
  "ggpubr",
  "spdep",
  "sfdep",
  "here"
)

# Install packages not yet installed
installed_packages <- packages %in% rownames(installed.packages())
if (any(installed_packages == FALSE)) {
  install.packages(packages[!installed_packages])
}

# Packages loading
invisible(lapply(packages, library, character.only = TRUE, quietly = TRUE))

library(sfislands)

```
```{r, echo=FALSE}
uk_election$country <- case_when(uk_election$region == "Wales" ~ "Wales",
                                  uk_election$region == "Scotland" ~ "Scotland",
                                  TRUE ~ "England")
```

## Workflow

The dataset `uk_elections`, which is provided with the package, is used for the following demonstrations. It is an `sf` dataset with geometries for each of the 632 Westminster parliamentary constituencies of England, Scotland and Wales in the 2019 general election. For each constituency, it contains socio-economic data from the 2011 census. It also shows the *Butler swing* between the 2017 and 2019 elections, which is a measure of the change in voter allegiance in favour of the Conservative Party from the Labour Party.

### Pre-functions: Creating desired adjacency structure

To demonstrate the `st_bridges()` function for preparing an adjacency matrix, we focus on Scotland. It is the nation with the highest number of islands. The function, with default arguments, returns a dataframe with an additional column named `nb`. This column is a named neighbourhood list of class `nb`. The default neighbourhood structure which it computes is:

1.    for constituencies which are **not islands**, first-order queen contiguity. A constituency is considered to be a neighbour of all of those with which it shares at least a vortex.

2.    for **islands**, contiguities are assigned to the two nearest constituencies to each island. This number can be altered using the `link_islands_k` argument.

```{r}
nbsf <- st_bridges(df = uk_election |> filter(country == "Scotland"),
                   geom_col_name = "constituency_name",
                   link_islands_k = 2)
```

This new dataframe has the following form (with a column named `nb` containing a neighbours list):

```{r}
head(nbsf)
```
We can inspect (and openly report) what additional contiguities have been added to the standard queen contiguity structure using the function `st_check_islands()`.

```{r}
st_check_islands(nbsf)
```

The overall contiguity structure can be visualised using the function `st_quickmap_nb()`, whose output is shown below on the left.

```{r, fig.width=12, fig.height=10}
ggarrange(
  st_quickmap_nb(nbsf),

  st_quickmap_nb(nbsf,
                 concavehull = TRUE,
                 nodes = "numeric",
                 numericsize = 3.5),

  ncol = 2
)

```

However, it is often the case that each *island* unit (in this case, constituency) is in fact a multipolygon of islands. To make this clearer, the argument `concavehull = TRUE` can be chosen. These green boundaries (shown above on the right) around each constituency are for reference purpose only and do not influence the assignment of contiguities. They show which group of *islands* are considered as one *constituency of islands*. Furthermore, the argument `nodes = "numeric"` represents the centroids of each unit with its index number instead of a geometric point. This can be useful for editing the adjacencies, particularly if the researcher is not familiar with the names of the individual spatial units.

#### Editing contiguities

A researcher can now use domain knowledge or a particular hypothesis to edit these contiguities. For instance, we may not be happy with the connection between 47 (Na h-Eileanan An Iar) and 51 (Orkney and Shetland) and wish to remove it. We might instead feel that it is important for 47 and 5 to be considered neighbours. This can be easily achieved, as shown below, using the `st_manual_cut_nb()` and `st_manual_join_nb()` functions. The spatial units whose relationships we wish to change can be referenced by name or by index number.

```{r}
nbsf <- st_bridges(df = uk_election |> filter(country == "Scotland"),
                   geom_col_name = "constituency_name",
                   link_islands_k = 2) |> 
  st_manual_cut_nb("Na h-Eileanan An Iar", "Orkney and Shetland") |> 
  st_manual_join_nb(47, 5)
```

The results of these changes as visualised in map form are shown below:

```{r, fig.width=12, fig.height=10}
ggarrange(
  st_quickmap_nb(nbsf,
                 concavehull = TRUE,
                 nodes = "numeric",
                 numericsize = 3.5),

  st_quickmap_nb(nbsf),

  ncol = 2
)

```

The above process can be iterated until we are happy that the structure conforms with our hypothesis of connectivity within the study area.

### Modelling & post-functions

Having set up a neighbourhood structure and embedded it as a named list or matrix within the original `sf` dataset as a column named `nb`, the workflow to generate and visualise predictions from a model (in this case using `mgcv`) is very straightforward and consists of only 3-4 lines of code.

For example, we can use the `mgcv` package to generate quite a complicated model with various random intercepts and slopes at region and county level, and additionally, different Markov Random Field ICAR smooths at the constituency level. This is done very quickly by

* using `st_bridges()` to prepare the data, 
* putting that inside the `mgcv::gam()` formulation, and then 
* piping into the `st_augment()` function.

This gives the following output, with all of the spatially varying predictions and their standard errors being added as extra columns before the final geometry column:

```{r}
prep_data <- st_bridges(uk_election, "constituency_name") # decide upon the contiguities and add them to the df

model1 <- gam(con_swing ~ 
               s(region, bs="re") + # region level random intercept
               s(county, bs="re") + # county level random intercept
               s(county, degree_educated, bs="re") + # county level random coefficient
               s(constituency_name, bs='mrf', 
                 xt=list(nb=prep_data$nb),k=10) + # ICAR constituency ICAR varying coefficients
               s(constituency_name, by=white, bs='mrf', 
                 xt=list(nb=prep_data$nb),k=10), # ICAR constituency ICAR varying coefficients
             data=prep_data, method="REML") |>  # pipe into function to get estimates
  st_augment(prep_data)

head(model1)

```

The predictions can be quickly visualised using `st_quickmap_preds()`. As this function generates a list of `ggplots`, they can be conveniently visualised using the `ggarrange()` function from the `ggpubr` package.

```{r, fig.width=12, fig.height=6}
plots <- st_quickmap_preds(model1)
ggarrange(plotlist = plots, legend = "none", nrow=1)
```

To emphasise the simplicity of the process, the a modified version of the following sequence of commands will generate a series of plots of all of the spatially varying elements of a model defined in `mgcv`:

```{r, eval=FALSE}
prep_data <- st_bridges(...)

gam(..., data = prep_data) |> 
  st_augment() |> 
  st_quickmap_preds()

```

It is also possible to conduct the same kind of augmenting and plotting of models generated using the `lme4` and `nlme` packages.

Further information and vignettes containing more detailed examples are available [here](https://horankev.github.io/sfislands/).